---
title: "Meta Wards"
output: html_notebook
---

# Processing the census 

The purpose of this file is to collect information, and data required for the Wards Model. We need:
1) origin destination data, 


```{r libs}
library(tidyverse)
library(rgdal)
library(leaflet)
library(maptools)
library(broom)
# set factors to false
options(stringsAsFactors = FALSE)


```





# Processing Origin-Destination data


We need to generate a commuter matrix from Ward to Ward. Data are available from ONS. Below we detail the procedure if we need to repreat. 



1. Download Ward lookups for Census Merged Wards to Original Wards from 

https://geoportal.statistics.gov.uk/datasets/ward-to-census-merged-ward-to-local-authority-district-december-2011-lookup-in-england-and-wales

2. Then download the lookup from Output Areas to Ward level. 

https://geoportal.statistics.gov.uk/datasets/output-area-to-ward-to-local-authority-district-december-2018-lookup-in-england-and-wales

3. Use left_join to combine them into a master lookup table. 


```{r process }

OA2Ward = read.csv('~/GitHub/MetaWards/data/2011/Output_Area_to_Ward_to_Local_Authority_District_December_2011_Lookup_in_England_and_Wales.csv') # maps OA to Ward data

wardlookup<-read.csv('~/GitHub/MetaWards/data/2011/WardsProcessing/Ward_Lookup.csv')

OA2Ward %>% left_join(.,wardlookup,by=c('WD11CD'='WD11CD'))->OA2WardLookup

write.csv(file="WardLookupMaster.csv",OA2WardLookup)

```

4. Download the bulk data for Output Area to Output Area commuter numbers from:
https://www.nomisweb.co.uk/census/2011/bulk/rOD1

at Output Area level. 

5. We need to aggregate up to Ward level, which we do below. 

```{r process origin-destination}

#4

OA2OAmovements = read.csv(file='data/2011/WardsProcessing/bulk/wf01bew_oa_v1.csv', header=F)# Output area to output area commuting. 

#5


OA2OAmovements %>% 
  inner_join(.,OA2WardLookup, by = c('V1'='OA11CD')) %>% # match output area  to ward for first column
  inner_join(.,OA2WardLookup, by = c('V2'='OA11CD'))%>% # match output area  to ward for second column
  select(from=FID.x,to=FID.y,ObjectID=ObjectId.x,V3) %>%        # remove columns  that are not needed
  group_by(from,to) %>% 
  summarize(total=sum(V3)) %>% 
  ungroup->Ward2Ward



```

The Ward identifier is "FID", which seems to go from 1 to 8588, and is the index which will be used in the code. 


```{r}
Ward2Ward %>% ggplot(aes(x=total)) +
  geom_freqpoly()+
  scale_x_log10()+
  scale_y_log10()+ # checking the histogram makes sense. 


write.table(file='EW.dat',Ward2Ward,col.names = FALSE,row.names=F)


```

We also need a population size per ward, population working per ward and population not working per ward. 



# Ward population sizes

Population sizes from here:
https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/wardlevelmidyearpopulationestimatesexperimental


```{r}

PopSizePerWard=read.csv('data/2011/PopSizePerWard.csv') # table from above

wardlookup %>% 
  inner_join(.,PopSizePerWard,by=c('WD11CD'='geography.code'))->wardlookupWithPop # join with population ward lookup tables

Ward2Ward %>% 
  group_by(from) %>% 
  summarise(WorkSize=sum(total)) ->WorkSize # sum up the populations commuting to give us the commuting size

write.table(file='WorkSize.dat',WorkSize,col.names = FALSE,row.names=F) # write WorkPopulation per ward to file

WorkSize %>% 
  inner_join(.,wardlookupWithPop,by=c("from"="FID"))->wardlookupWithPop2 # join with ward lookup file

wardlookupWithPop2 %>% 
  select(Ward=from,All=Variable..All.usual.residents..measures..Value,WorkSize) %>% 
  mutate(PlaySize=All-WorkSize) %>% 
  mutate(PlaySize=ifelse(PlaySize<0,0,PlaySize))->wardlookupWithPop3

wardlookupWithPop3%>% 
  select(from=Ward,PlaySize) %>% 
  write.table(.,file="PlaySize.dat",col.name=FALSE,row.names=FALSE)

```

# Non-commuter matrix

```{r}
wardlookupWithPop3%>% 
  select(from=Ward,PlaySize)-> 
  play


Ward2Ward %>% 
  group_by(from) %>% 
  mutate(rate=total/sum(total)) %>% 
  ungroup() %>% 
  inner_join(.,play,by='from') %>% 
  mutate(PlayTotal=round(rate*PlaySize))->Ward2WardAll
  
Ward2WardAll %>% 
  select(from,to,rate) %>% 
  write.table(.,file="PlayMatrix.dat",row.names=F,col.names=F)


```

# CCG analysis

```{r}
lsoa2ccg = read.csv('~/GitHub/MetaWards/data/2011/CCG/Lower_Layer_Super_Output_Area_2011_to_Clinical_Commissioning_Group_to_Local_Authority_District_April_2017_Lookup_in_England_Version_4.csv')

lsoa2ccg %>% 
  group_by(CCG17CD) %>% 
  count()



```

